---
title: "CopyKAT Analysis Report"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: cosmo
    code_folding: hide
params:
  sample_name: "sample"
  output_dir: "."
  result: NULL
  config: NULL
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

# Load required libraries
library(knitr)
library(ggplot2)
```

# Analysis Overview

**Sample Name:** `r params$sample_name`

**Analysis Date:** `r format(Sys.time(), '%B %d, %Y at %I:%M %p')`

**Output Directory:** `r params$output_dir`

---

## Summary Statistics

TODO: Display summary statistics

```{r summary-stats}
# TODO: Extract and display summary statistics from params$result
# - Total cells analyzed
# - Number of aneuploid cells
# - Number of diploid cells
# - Fraction of aneuploid cells
```

---

## Parameters Used

TODO: Display analysis parameters

```{r parameters}
# TODO: Display parameters from params$config
# Create a table showing all CopyKAT parameters used
```

---

## Cell Classifications

TODO: Display cell classification results

```{r cell-classifications}
# TODO: Read and display predictions file
# predictions_file <- file.path(params$output_dir, 
#                               paste0(params$sample_name, "_copykat_prediction.txt"))
```

---

## CNV Heatmap

TODO: Display CNV heatmap

```{r cnv-heatmap, fig.width=10, fig.height=8}
# TODO: Display the CNV heatmap image
# heatmap_file <- file.path(params$output_dir,
#                          paste0(params$sample_name, "_copykat_heatmap.jpeg"))
# knitr::include_graphics(heatmap_file)
```

---

## Classification Distribution

TODO: Plot classification distribution

```{r classification-plot, fig.width=8, fig.height=6}
# TODO: Create bar plot of cell type classifications
```

---

## Conclusions

TODO: Add interpretation and conclusions

---

## Session Information

```{r session-info}
sessionInfo()
```

---

**Report generated by CopyKAT Analysis Pipeline**

**NOTE TO BACKEND TEAM:** This is a skeleton R Markdown template.
Complete implementation available in the original copykat_report.Rmd.
Fill in TODO sections with actual analysis results and visualizations.

